---
title: "cyanosphere_pcoa_update_18Jan2022"
author: "Brianne"
date: "1/18/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/Brianne/Documents/Cyanosphere/")
set.seed(12345)
```

# Load Libraries
```{r}
library(ecodist)
library(tidyverse)
library(vegan)
library(viridis)
library(ggrepel)
library(asbio)
```

# load data
```{r}
cyanosphere <- read.csv("data/autometa_assembly_cyanosphere.csv")
cyanosphere.transformed <- cyanosphere %>% pivot_wider(names_from = taxid, values_from = contigs)
cyano.all <-cyanosphere.transformed
```

Make all values numeric
```{r}
cyano.all[c(51:349)] <- sapply(cyano.all[c(51:349)],as.numeric)
cyano.all[c(51:349)][is.na(cyano.all[c(51:349)])] <-0
```

Group by culture_ID to get just 50 points 

Group dataframe by culture ID and maintain metadata
```{r}
otu.all <- cyano.all %>% group_by(culture_ID, Habitat, Location, Continent, Substrate, Host_Phylum, Host_Class, Host_Order, Host_Family, Host_Genus) %>% summarize_at(vars(bio1:`159191`),list(mean = mean))
```


# create presence absence matrix 
```{r}
otu.all[c(30:328)] <- ifelse(otu.all[c(30:328)]>0, 1, 0)
otu.only <- otu.all[c(30:328)]
metadata <- otu.all[c(1:29)]
```

Calculate beta diversity
beta = gamma/alpha -1

```{r}
alpha <- alpha.div(otu.only, "simp")
gamma <- with(metadata, specnumber(otu.only, culture_ID))

beta <- as.matrix(gamma/alpha -1)
```

```{r}
metadata$beta <- beta
beta1 <- beta[-34,]
```

Anova on beta diversity 

```{r}
adonis(beta1 ~ culture_ID, metadata[-34,])
```

# make bray-curtis matrix
```{r}
otu.bray <- vegdist(otu.only, method = "bray")# dissimilarity matrix using bray-curtis distance indices 
```

```{r}
pcoaVS <- pco(otu.bray, negvals = "zero", dround = 0) # if negvals = 0 sets all negative eigenvalues to zero
```

Look at environmental variables 
```{r}
env <- metadata[c(2:29)]

# CMD
cmd <- isomapdist(vegdist(otu.only, method="bray"), k=3) #cmdscale is classical metric multidimensional scaling and this is the way to measure the effect of environmetnal variables on the community composition, this analysis doesn't work with the pcoa created earlier
```

```{r}
# Plot CMD
ordiplot(cmd) #plot looks similar to pcoa
```

Make dataframe for plot with vectors and metadata
```{r}
pcoa1 = pcoaVS$vectors[,1] # adds the points of the NMDS 1 dimmension
pcoa2 = pcoaVS$vectors[,2] # adds the points of the NMDS 2 dimmension

MDS = data.frame(pcoa1 = pcoa1, pcoa2 = pcoa2)
MDS = cbind(MDS, metadata)
```

Make plot
```{r}
ggplot(MDS, aes(x = pcoa1, y = pcoa2)) + geom_point(aes(color = culture_ID), size = 2) + 
  theme_bw() + theme(legend.position = "none") + geom_label(data = MDS, aes(label = culture_ID))

```

```{r}
# Fit environmental variables
set.seed(12345)
fit_env <- envfit(ord=cmd, env=env) 
fit_env$vectors
```

```{r}
arrow <- data.frame(fit_env$vectors$arrows, R = fit_env$vectors$r, P = fit_env$vectors$pvals) # make dataframe with vectors for arrows
arrow$FG <-  rownames(arrow)
arrow.p <- filter(arrow, P<=0.05)
arrow.p$label <- c("Annual Mean Temp", "Min Temp Coldest Month", "Mean Temp Coldest Quarter", "Annual Precip","Precip Wettest Month", "Precip Driest Month", "Precip Wettest Quarter", "Precip Driest Quarter", "Precip Warmest Quarter","Precip Coldest Quarter")
```

```{r}
pdf("figures/PCOA_cultureID_envArrows.pdf")
ggplot(data=MDS, aes(x = pcoa1, y = pcoa2)) + geom_point(aes(color = Substrate)) + theme_minimal() + geom_segment(data=arrow.p, aes(x=0, y=0, xend= X1, yend=X2), arrow=arrow(length=unit(0.001, "mm")*arrow.p$R)) + geom_label_repel(data=arrow.p, aes(x=X1, y=X2,  label = label),  label.padding = 0.1, label.size = 0.1, size =3, max.overlaps = 30)
dev.off()
```
Permanova on all variables
```{r}
set.seed(12345)
adonis(otu.only ~ Habitat, metadata) # p = 0.011
adonis(otu.only ~ Location, metadata) # p = 0.001
adonis(otu.only ~ Continent, metadata) # p = 0.007
adonis(otu.only ~ Substrate, metadata) # p = 0.018
adonis(otu.only ~ Host_Class, metadata) # p = 0.65
adonis(otu.only ~ Host_Order, metadata) # p = 0.026
adonis(otu.only ~ Host_Family, metadata) # p = 0.007
adonis(otu.only ~ Host_Genus, metadata) # p = 0.018
```

```{r}
library(pairwiseAdonis)
pairwise.adonis(otu.only,metadata$Habitat)
```

```{r}
ggplot(MDS, aes(x = pcoa1, y = pcoa2)) + geom_point(aes(color = Substrate), size = 3, position=position_jitter(width=.1,height=.1)) + 
  theme_bw() + theme(legend.position = "right") + stat_ellipse(aes(color = Substrate))

```

```{r}
ggplot(MDS, aes(x = pcoa1, y = pcoa2)) + geom_point(aes(color = Habitat), size = 3, position=position_jitter(width=.1,height=.1)) + 
  theme_bw() + theme(legend.position = "right") + stat_ellipse(aes(color = Habitat))

```

```{r}
ggplot(MDS, aes(x = pcoa1, y = pcoa2)) + geom_point(aes(color = Location), size = 3, position=position_jitter(width=.1,height=.1)) + 
  theme_bw() + theme(legend.position = "right") + stat_ellipse(aes(color = Location))

```

```{r}
ggplot(MDS, aes(x = pcoa1, y = pcoa2)) + geom_point(aes(color = Continent), size = 3, position=position_jitter(width=.1,height=.1)) + 
  theme_bw() + theme(legend.position = "right") + stat_ellipse(aes(color = Continent))

```

```{r}
ggplot(MDS, aes(x = pcoa1, y = pcoa2)) + geom_point(aes(color = Host_Order), size = 3, position=position_jitter(width=.1,height=.1)) + 
  theme_bw() + theme(legend.position = "right") + stat_ellipse(aes(color = Host_Order))

```

```{r}
ggplot(MDS, aes(x = pcoa1, y = pcoa2)) + geom_point(aes(color = Host_Family), size = 3, position=position_jitter(width=.1,height=.1)) + 
  theme_bw() + theme(legend.position = "right") + stat_ellipse(aes(color = Host_Family))

```

```{r}
ggplot(MDS, aes(x = pcoa1, y = pcoa2)) + geom_point(aes(color = Host_Genus), size = 3, position=position_jitter(width=.1,height=.1)) + 
  theme_bw() + theme(legend.position = "right") + stat_ellipse(aes(color = Host_Genus))

```


