---
title: "Untitled"
author: "Brianne"
date: "12/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/Brianne/Documents/Cyanosphere/")
```

# Load Libraries
```{r}
library(ecodist)
library(tidyverse)
```
# load data
```{r}
cyanosphere <- read.csv("data/autometa_assembly_cyanosphere.csv")
cyanosphere.transformed <- cyanosphere %>% pivot_wider(names_from = taxid, values_from = contigs)
cyano.all <-cyanosphere.transformed
```

Make all values numeric
```{r}
cyano.all[c(50:348)] <- sapply(cyano.all[c(50:348)],as.numeric)
cyano.all[is.na(cyano.all)] <-0
```

Group dataframe by culture ID, habitat, location, continent, substrate, and host phylogeny 
```{r}
otu.all <- cyano.all %>% group_by(culture_ID, Habitat, Location, Continent, Substrate, Host_Phylum, Host_Class, Host_Order, Host_Family, Host_Genus) %>% summarize_at(vars(bio1:`159191`),list(mean = mean))
```

# create presence absence matrix 
```{r}
otu.all[c(30:328)] <- ifelse(otu.all[c(30:328)]>0, 1, 0)
```

Make numeric matrix and metadata matrix
```{r}
otu.only <- otu.all[c(30:328)]
head(otu.only)
metadata <- otu.all[-c(30:328)]
```

# make bray-curtis matrix
```{r}
otu.bray <- vegdist(otu.only, method = "bray") # dissimilarity matrix using bray-curtis distance indices 
```

```{r}
pcoaVS <- pco(otu.bray, negvals = "zero", dround = 0) # if negvals = 0 sets all negative eigenvalues to zero
```

```{r}
pcoaVS$values # eigenvalue for each component. This is a measure of the variance explained by each dimension
pcoaVS$vectors # eigenvectors. Each column contains the scores for that dimension.
```

Make dataframe for plot with vectors and metadata
```{r}
pcoa1 = pcoaVS$vectors[,1] # adds the points of the NMDS 1 dimmension
pcoa2 = pcoaVS$vectors[,2] # adds the points of the NMDS 2 dimmension

MDS = data.frame(pcoa1 = pcoa1, pcoa2 = pcoa2)
MDS = cbind(MDS, metadata)
```

Make plot
```{r}
ggplot(MDS, aes(x = pcoa1, y = pcoa2)) + geom_point(aes(color = Substrate), size = 2) + 
  theme_bw() + theme(legend.position = "right") #+ stat_ellipse(aes(color = Host_Family))

```

PERMANOVA
```{r}
adonis(otu.only ~ Substrate, metadata)
```

# Permanova on all variables
```{r}
adonis(otu.only ~ culture_ID, metadata) # p=1
adonis(otu.only ~ Habitat, metadata) # p = 0.01
adonis(otu.only ~ Location, metadata) # p = 0.002
adonis(otu.only ~ Continent, metadata) # p = 0.008
adonis(otu.only ~ Substrate, metadata) # p = 0.05
adonis(otu.only ~ Host_Class, metadata) # p = 0.634
adonis(otu.only ~ Host_Order, metadata) # p = 0.042
adonis(otu.only ~ Host_Family, metadata) # p = 0.005
adonis(otu.only ~ Host_Genus, metadata) # p = 0.018
```
Look at environmental variables 
```{r}
env <- metadata[c(11:29)]

# CMD
cmd <- isomapdist(vegdist(otu.only, method="bray")) #cmdscale is classical metric multidimensional scaling and this is the way to measure the effect of environmetnal variables on the community composition, this analysis doesn't work with the pcoa created earlier
```


```{r}
# Plot CMD
ordiplot(cmd) #plot looks similar to pcoa
```

```{r}
# Fit environmental variables
fit_env <- envfit(ord=cmd, env=env) 
fit_env$vectors
```

```{r}
arrow <- data.frame(fit_env$vectors$arrows, R = fit_env$vectors$r, P = fit_env$vectors$pvals) # make dataframe with vectors for arrows
arrow$FG <-  rownames(arrow)
arrow.p <- filter(arrow, P<=0.05)
```

```{r}
ggplot(data=MDS, aes(x = pcoa1, y = pcoa2)) + geom_point(data=MDS,aes(color = Substrate), position=position_jitter(.1)) + theme_minimal() + geom_segment(data=arrow.p, aes(x=0, y=0, xend= Dim1, yend=Dim2), arrow=arrow(length=unit(.2, "cm")*arrow.p$R)) + geom_label_repel(data=arrow.p, aes(x=Dim1, y=Dim2,  label = FG),  label.padding = 0.1, label.size = 0.1, size =3, max.overlaps = 30)
```
