---
title: "Heat Maps"
author: "Brianne"
date: "12/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/Brianne/Documents/Cyanosphere/")
```

# what organisms are present in the cyanosphere 

# load libraries 
```{r}
library(tidyverse)
library(vegan)
library(ggrepel)
```

# load data
```{r}
cyanosphere <- read.csv("data/autometa_assembly_cyanosphere.csv")
```

```{r}
ggplot(cyanosphere, aes(fill = phylum, y = contigs, x = culture_ID)) + geom_bar(position = "fill", stat = "identity")+ theme_bw()  + labs(y = "% abundance") + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1), text = element_text(size = 8), plot.margin = unit(c(0.2, 0.2, 0.2, 2), "cm"))
```

Make Heatmaps

```{r}
heatmap.data <- cyanosphere %>% dplyr::select(taxid, contigs, culture_ID)
heatmap.data <- pivot_wider(heatmap.data, names_from = taxid, values_from = contigs, values_fn = sum)
culture.names <-  heatmap.data$culture_ID
heatmap.data <- heatmap.data[-1]
heatmap.data[c(1:299)] <- ifelse(heatmap.data[c(1:299)]>0, 1, 0)
heatmap.data <- sapply(heatmap.data, as.numeric)
heatmap.data[is.na(heatmap.data)] <-0
rownames(heatmap.data) <- culture.names

heatmap(heatmap.data, margins = c(10, 11.5))
```
Heat nap showing the abundance of different phyla in each culture
```{r}
heatmap.data <- cyanosphere %>% dplyr::select(phylum, contigs, culture_ID)
heatmap.data <- pivot_wider(heatmap.data, names_from = phylum, values_from = contigs, values_fn = sum)
culture.names <-  heatmap.data$culture_ID
heatmap.data <- heatmap.data[-1]
heatmap.data[c(1:12)] <- ifelse(heatmap.data[c(1:12)]>0, 1, 0)
heatmap.data <- sapply(heatmap.data, as.numeric)
heatmap.data[is.na(heatmap.data)] <-0
rownames(heatmap.data) <- culture.names

heatmap(heatmap.data, margins = c(10, 11.5))
```

look just at proteobacteria 

```{r}
proteo <- cyanosphere %>% filter(phylum == "Proteobacteria")

ggplot(proteo, aes(fill = order, y = contigs, x = culture_ID)) + geom_bar(position = "fill", stat = "identity")+ theme_bw()  + labs(y = "% abundance") + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1), text = element_text(size = 8), plot.margin = unit(c(0.2, 0.2, 0.2, 2), "cm"))
```

proteobacteria orders in each culture
```{r}
heatmap.data <- proteo %>% dplyr::select(order, contigs, culture_ID)
heatmap.data <- pivot_wider(heatmap.data, names_from = order, values_from = contigs, values_fn = sum)
culture.names <-  heatmap.data$culture_ID
heatmap.data <- heatmap.data[-1]
heatmap.data[c(1:12)] <- ifelse(heatmap.data[c(1:12)]>0, 1, 0)
heatmap.data <- sapply(heatmap.data, as.numeric)
heatmap.data[is.na(heatmap.data)] <-0
rownames(heatmap.data) <- culture.names

heatmap(heatmap.data, margins = c(10, 11.5))
```


Based on Substrate

```{r}

ggplot(cyanosphere, aes(fill = phylum, y = contigs, x = Substrate)) + geom_bar(position = "fill", stat = "identity")+ theme_bw()  + labs(y = "% abundance") + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1), text = element_text(size = 8), plot.margin = unit(c(0.2, 0.2, 0.2, 2), "cm"))
```

What phyla are present in each substrate?
```{r}
heatmap.data <- cyanosphere %>% dplyr::select(phylum, contigs, Substrate)
heatmap.data <- pivot_wider(heatmap.data, names_from = phylum, values_from = contigs, values_fn = sum)
substrate.names <-  heatmap.data$Substrate
heatmap.data <- heatmap.data[-1]
heatmap.data[c(1:12)] <- ifelse(heatmap.data[c(1:12)]>0, 1, 0)
heatmap.data <- sapply(heatmap.data, as.numeric)
heatmap.data[is.na(heatmap.data)] <-0
rownames(heatmap.data) <- substrate.names

heatmap(heatmap.data, margins = c(10, 11.5))

```
proteobacteria orders in each substrate
```{r}
heatmap.data <- proteo %>% dplyr::select(order, contigs, Substrate)
heatmap.data <- pivot_wider(heatmap.data, names_from = order, values_from = contigs, values_fn = sum)
heatmap.data <- heatmap.data[-1]
heatmap.data[c(1:12)] <- ifelse(heatmap.data[c(1:12)]>0, 1, 0)
heatmap.data <- sapply(heatmap.data, as.numeric)
heatmap.data[is.na(heatmap.data)] <-0
rownames(heatmap.data) <- substrate.names

heatmap(heatmap.data, margins = c(10, 11.5))

```

Based on Continent

```{r}
ggplot(cyanosphere, aes(fill = phylum, y = contigs, x = Continent)) + geom_bar(position = "fill", stat = "identity")+ theme_bw()  + labs(y = "% abundance") + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1), text = element_text(size = 8), plot.margin = unit(c(0.2, 0.2, 0.2, 2), "cm"))
`dev.off()
```
whay phyla are present on each continent?
```{r}
heatmap.data <- cyanosphere %>% dplyr::select(phylum, contigs, Continent)
heatmap.data <- pivot_wider(heatmap.data, names_from = phylum, values_from = contigs, values_fn = sum)
row.names <-  heatmap.data$Continent 
heatmap.data <- heatmap.data[-1]
heatmap.data[c(1:12)] <- ifelse(heatmap.data[c(1:12)]>0, 1, 0)
heatmap.data <- sapply(heatmap.data, as.numeric)
heatmap.data[is.na(heatmap.data)] <-0
rownames(heatmap.data) <- row.names

heatmap(heatmap.data, margins = c(10, 11.5))
```

Proeteobacteria only 

```{r}
heatmap.data <- cyanosphere %>% filter(phylum == "Proteobacteria") %>% dplyr::select(order, contigs, Continent)
heatmap.data <- pivot_wider(heatmap.data, names_from = order, values_from = contigs, values_fn = sum)
row.names <-  heatmap.data$Continent 
heatmap.data <- heatmap.data[-1]
heatmap.data[c(1:12)] <- ifelse(heatmap.data[c(1:12)]>0, 1, 0)
heatmap.data <- sapply(heatmap.data, as.numeric)
heatmap.data[is.na(heatmap.data)] <-0
rownames(heatmap.data) <- row.names

heatmap(heatmap.data, margins = c(10, 11.5))
`dev.off()
```

based on Host Order 

```{r}
heatmap.data <- cyanosphere %>% dplyr::select(phylum, contigs, Host_Order)
heatmap.data <- pivot_wider(heatmap.data, names_from = phylum, values_from = contigs, values_fn = sum)
row.names <-  heatmap.data$Host_Order 
heatmap.data <- heatmap.data[-1]
heatmap.data[c(1:12)] <- ifelse(heatmap.data[c(1:12)]>0, 1, 0)
heatmap.data <- sapply(heatmap.data, as.numeric)
heatmap.data[is.na(heatmap.data)] <-0
rownames(heatmap.data) <- row.names

heatmap(heatmap.data, margins = c(10, 11.5))
```
Host order and proteobacteria
```{r}
heatmap.data <- cyanosphere %>% filter(phylum == "Proteobacteria") %>% dplyr::select(order, contigs, Host_Order)
heatmap.data <- pivot_wider(heatmap.data, names_from = order, values_from = contigs, values_fn = sum)
row.names <-  heatmap.data$Host_Order 
heatmap.data <- heatmap.data[-1]
heatmap.data[c(1:12)] <- ifelse(heatmap.data[c(1:12)]>0, 1, 0)
heatmap.data <- sapply(heatmap.data, as.numeric)
heatmap.data[is.na(heatmap.data)] <-0
rownames(heatmap.data) <- row.names

heatmap(heatmap.data, margins = c(12, 11.5))
```

